Equalizer: A Mature Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
co::Object section
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@11 db7c848b-7755-4df2-b763-8796797afa52

@@

types (instance, delta, static), versioning, multicast, compression,
serializable with dirty bits, mapping, blocking commits

\subsection{Reliable Stream Protocol}

reverse-engr impl

>>>>

An important part of writing a parallel rendering application is the
communication layer between the individual processes. Equalizer relies on the
Collage network library for its internal operation. Collage furthermore provides
powerful abstractions for writing Equalizer applications, which are introduced
in this section.

\subsection{Architecture}

Collage provides networking functionality of different abstraction layers,
gradually providing higher level functionality for the programmer. The main
primitives in Collage are:

\begin{compactdesc}
\item[Connection] A stream-oriented point-to-point communication
  line. The connections
  transmit raw data reliably between two endpoints for unicast connections, and
  between a set of endpoints for multicast connections. For unicast,
  process-local pipes, TCP and Infiniband RDMA are implemented. For multicast,
  a reliable, UDP-based protocol is discussed in \sref{SEC_RSP}.
\item[DataI/OStream] Abstracts the input and output of C++ data types from or to
  a set of connections by implementing output stream operators. Uses buffering
  to aggregate data for network transmission. Performs endian swapping during
  input if the endianness differs between the originating and local node.
\item[Node and LocalNode] The abstraction of a process in the cluster. Nodes
  communicate with each other using connections. A LocalNode listens on various
  connections and processes requests for a given process. Received data is
  wrapped in ICommands and dispatched to command handler methods. A Node is a
  proxy for a remote LocalNode. The Equalizer Client object is a LocalNode.
\item[Object] Provides object-oriented, versioned data distribution of C++
  objects between nodes. Objects are registered or mapped on a Local\-Node.
\end{compactdesc}

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{EqDocs/Developer/ProgrammingGuide/images/netObject}
  \caption{\label{fNetObject}Communication between two Collage Objects}
\end{figure}
\fig{fNetObject} illustrates this architecture.

needed? reverse-engr impl

\subsection{Reliable Stream Protocol}\label{SEC_RSP}

@@

Adapting an existing application for parallel rendering requires the
synchronization of application data across the processes in the parallel
rendering setup. Existing parallel rendering frameworks address this often
poorly, at best they rely on MPI to distribute data. Real-world, interactive
visualization applications are typically written in C++ and have complex data
models and class hierarchies to represent their application state. As outlined
in \cite{EMP:09}, the parallel rendering code in an Equalizer application only
needs access to the data needed for rendering, as all application logic is
centralized in the application main thread. We have encountered two main
approaches to address this distribution: Using a shared filesystem for static
data and using data distribution for static and dynamic data.

Distributed objects in Collage provide powerful, object-oriented data
distribution for C++ objects. They facilitate the implementation of data
distribution in a cluster environment. Distributed objects are created by
subclassing from \textsf{co::Serializable} or \textsf{co::Object}. The
application programmer implements serialization and deserialization of the
distributed data. Distributed objects can be static (immutable) or
dynamic. Objects have a universally unique identifier (UUID) to address them
cluster-wide. A master-slave model is used to establish mapping and data
synchronization across processes. Typically, the application mail loop registers
a master instance and communicates the UUID to the render clients, which map
their instance to the given identifier. The following object types are
available:

\begin{compactdesc}
\item[Static] The object is not versioned nor buffered. The instance data is
  serialized whenever a new slave instance is mapped. No additional data is
  stored.
\item[Instance] The object is versioned and buffered. The instance and delta
  data are identical, that is, only instance data is serialized. Previous
  instance data is saved to be able to map old versions.
\item[Delta] The object is versioned and buffered. The delta data is typically
  smaller than the instance data. The delta data is transmitted to slave
  instances for synchronization. Previous instance and delta data is saved to be
  able to map and sync old versions.
\item[Unbuffered] The object is versioned and unbuffered. No data is stored, and
  no previous versions can be mapped. No old versions can be mapped and no
  additional data is stored.
\end{compactdesc}

All distributed objects have to implement \textsf{getInstanceData} and
\textsf{applyInstanceData} to serialize and deserialize the object's distributed
data. These methods provide an output or input stream as a parameter, which
abstracts the data transmission and can be used like a \textsf{std::stream}. The
data streams implement efficient buffering and compression, and automatically
select the best connection for data transport. Custom data type serializers can
be implemented by providing the appropriate serialization functions. No pointers
should be directly transmitted through the data streams. For pointers, the
corresponding object is typically a distributed object as well, and its
identifier and potentially version is transmitted in place of the memory
address.

Dynamic objects are versioned, and have to override \textsf{getChangeType} to
indicate how they want to have changes to be handled. Upon \textsf{commit} the
delta data from the previous version is sent, if available using multicast, to
all mapped slave instances. The data is queued on the remote node, and is
applied when the application calls \textsf{sync} to synchronize the object to a
new version. The \textsf{sync} method might block if a version has not yet been
committed or is still in transmission. All versioned objects the following
characteristics:

\begin{compactitem}
\item The master instance of the object generates new versions for all
  slaves. These versions are continuous. It is possible to commit on slave
  instances, but special care has to be taken to handle possible
  conflicts.
\item Slave instance versions can only be advanced, that is, \textsf{sync(
  version )} with a version smaller than the current version will fail.
\item Newly mapped slave instances are mapped to the oldest available
  version by default, or to the version specified when calling
  \textsf{mapObject}.
\end{compactitem}

Besides the instance data serialization methods used to map an object, versioned
objects may implement \textsf{pack} and \textsf{unpack} to serialize or
deserialize the changes since the last version.

The Collage Serializable implements one convenient usage pattern for object data
distribution. The \textsf{co::Serializable} data distribution is based on the
concept of dirty bits, allowing inheritance with data distribution. Dirty bits
form a 64-bit mask which marks the parts of the object to be distributed during
the next commit. For serialization, the application developer implements
\textsf{serialize} or \textsf{deserialize}, which are called with the bit mask
specifying which data has to be transmitted or received. During a commit or
sync, the current dirty bits are given, whereas during object mapping all dirty
bits are passed to the serialization methods.

Blocking commits allow to limit the number of outstanding, queued versions on
the slave nodes. A token-based protocol will block the commit on the master
instance if to many unsynchronized versions exist.
